---
title: "Exercise 5"
author: "Claire Brandes"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: leonids
    toc: yes
---

# Challenge 1

## Step 1

Using the {tidyverse} read_csv() function, load the "IMDB-movies.csv" data set as a "tibble" named **d**. 

```{r}
library(readr)
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
```

## Step 2

Use a one-line statement to filter the data set to include just movies from 1920 to 1979 and moves that are between 1 and 3 hours long (**runtimeMinutes** >= 60 and **runtimeMinutes** <= 180), and add a new column that codes the startYear into a new variable, **decade** ("20s", "30s",..."70s"). There are now 5651 movies remaining in the data set. 

```{r}
library(dplyr)
d <- d %>%
  filter(between(startYear, 1920, 1979)) %>%
  filter(between(runtimeMinutes, 60, 180)) %>%
  mutate(decade = case_when(
    startYear <= 1929 ~ "20s",
    startYear <= 1939 ~ "30s",
    startYear <= 1949 ~ "40s",
    startYear <= 1959 ~ "50s",
    startYear <= 1969 ~ "60s",
    startYear <= 1979 ~ "70s",
    TRUE ~ NA
  ))
```

## Step 3 

Use {ggplot2} to plot histograms of the distribution of **runtimeMinutes** for each decade. 

```{r}
library(tidyverse)
ggplot(data = d, mapping = aes(x = runtimeMinutes, group = runtimeMinutes, fill = runtimeMinutes)) + 
  ylab("Count") + xlab("Run-Time in Minutes") +
  geom_histogram(bins = 30) + 
  facet_wrap(vars(decade), nrow = 5, ncol = 2) + 
  theme_dark() 
```

## Step 4

Use a one-line statement to calculate the population mean and population standard deviation in runtimeMinutes for each decade and save the results in a new data frame called **results**. 


```{r}
(results <- d %>% group_by(decade) %>%
  summarise(
    pop_mean = mean(runtimeMinutes),
    pop_sd = sd(runtimeMinutes)
  ))
```

## Step 5

Draw a single sample of 100 movies, without replacement, from each decade and calculate the single sample mean and single sample standard deviation in **runtimeMinutes** for each decade.

```{r}
s <- d %>% group_by(decade) %>%
  sample_n(size = 100, replace = FALSE)
```

Calculating single sample mean and single standard deviation in runtimeMinutes for each decade:

```{r}
(sample_results <- s %>% group_by(decade) %>%
  summarise(
    sample_mean = mean(runtimeMinutes),
    sample_sd = sd(runtimeMinutes)
  ))
```

## Step 6

Calculate for each decade the standard error around your estimate of the population mean **runtimeMinutes** based on the standard deviation and sample size (n = 100 movies) of your single sample. Adding this data to the tibble created in Step 5:

```{r}
(sample_results <- s %>% group_by(decade) %>%
  summarise(
    sample_mean = mean(runtimeMinutes),
    sample_sd = sd(runtimeMinutes),
    std_error = sample_sd/sqrt(100)
  ))
```

## Step 7

Compare these estimates to the actual population mean **runtimeMinutes** for each decade and to the calculated SE in the population mean for samples of size 100 based on the population standard deviation for each decade. 

```{r}
full_join(results, sample_results)
```

## Step 8

Generate a *sampling distribution* of mean **runtimeMinutes** for each decade by [a] drawing 1000 random samples of 100 movies from each decade, without replacement, and, for each sample, [b] calculating the mean **runtimeMinutes** and the standard deviation in **runtimeMinutes** for each decade. 

Generating sampling distribution:

```{r}
library(mosaic)
reps <- 1000

samp_dist <- do(reps) * sample_n(group_by(d, decade), 100, replace = FALSE)
```

Calculating mean and std. dev. for each decade:

```{r}
(samp_dist_results <- samp_dist %>% group_by(decade) %>%
  summarise(
    samp_dist_mean = mean(runtimeMinutes),
    samp_dist_sd = sd(runtimeMinutes) 
  ))
```

## Step 9

Calculate the **mean** and the **standard deviation** of the sampling distribution of sample means for each decade and plot a histogram of the sampling distribution for each decade.

Calculating mean and std. dev. of sampling distribution:

```{r}
mean(samp_dist_results$samp_dist_mean)
sd(samp_dist_results$samp_dist_sd)
```

Plotting histogram of the sampling distribution for each decade:

```{r}

```

